### 2021-09-02 A methodological framework to enable the roaming agreement digitalization: An NLP-based approach/sections/Results.tex
01.08.21.4

\section{Discussion of Results}\label{results}
Our NLP Engine architecture receives PDF files as inputs, and PDF files normally consist of unstructured text. Because of this, undesired characters may remain despite text parsing. Therefore, it is mandatory to determine the \textit{accuracy} of the results obtained once the output JSON file has been populated. For this purpose, two types of experiments have been conducted to determine the \textit{accuracy} of the results. On the one hand, a simple inspection at the sub-article level and on the other hand, a verification based on symbol comparison. The tests have been performed on two real RA samples from the mobile operators Proximus and Orange \cite{proximus}. Each experiment is described below and the results obtained are discussed in the same section.

\subsection{Accuracy determination based on a simple inspection at sub-article level}
The first accuracy analysis consists of visually determining, i.e., human-eye inspection, whether each sub-article of the RA constitutes a \textit{variation}, a \textit{standard clause} or a \textit{customized text} concerning the GSMA standard template. The results obtained are then compared with the values populated in the NLP Engine output file, for the same sub-article. Considering that the results obtained from the simple inspection constitute the observations and the values collected from the sub-articles classification file constitute the predicted values, the following confusion matrices allow us to analyze the results for each sample RA. In this regard, the confusion matrix in Table~\ref{table1}  shows satisfactory results for the NLP Engine's predictive capability, i.e., regarding the accuracy at which the NLP Engine correctly classifies 61 of 72 analyzed sub-articles as \textit{standard clauses}, \textit{variations} or \textit{customized texts} for the Proximus sample RA. Proof of this is the fact that the main diagonal of the matrix contains the highest number of True Positives. Table~\ref{table3} summarizes the accuracy of the NLP Engine's sub-article classification capability in percentage terms as \textit{standard clauses}, \textit{variations} and \textit{customized texts}.

\begin{table}[htbp]
\caption{Confusion Matrix for Proximus Roaming Agreement.}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{n = 72} & \textbf{\textit{stdClause}}& \textbf{\textit{variation}}& \textbf{\textit{customText}} \\
\hline
\textbf{\textit{stdClause}}& 21 & 3 & 1 \\
\hline
\textbf{\textit{variation}}& 2 & 35 & 4 \\
\hline
\textbf{\textit{customText}}& 0 & 1 & 5 \\
\hline
\end{tabular}
\label{table1}
\end{center}
\end{table}

\begin{table}[htbp]
\caption{Confusion Matrix for Orange Roaming Agreement.}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{n = 86} & \textbf{\textit{stdClause}}& \textbf{\textit{variation}}& \textbf{\textit{customText}} \\
\hline
\textbf{\textit{stdClause}}& 14 & 0 & 1 \\
\hline
\textbf{\textit{variation}}& 1 & 8 & 0 \\
\hline
\textbf{\textit{customText}}& 4 & 1 & 57 \\
\hline
\end{tabular}
\label{table2}
\end{center}
\end{table}

Table~\ref{table2} shows satisfactory results for the NLP Engine's predictive capability, i.e., regarding the \textit{accuracy} at which the NLP Engine correctly classifies 79 of the 86 sub-articles as \textit{standard clauses}, \textit{variations}, or \textit{customized texts} for the Orange sample RA. Using the same evaluation criteria, it is observed how the main diagonal of the confusion matrix in Table~\ref{table2} has the highest number of True Positive. Table~\ref{table3} also summarizes the accuracy of the NLP Engine's sub-article classification capability in percentage terms as \textit{standard clauses}, \textit{variations} and \textit{customized texts}.

\begin{table}[htbp]
\caption{Summary of accuracy determination for simple inspection at sub-article level.}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{} & \textbf{\textit{Proximus}}& \textbf{\textit{Orange}} \\
\hline
\textbf{\textit{stdClause}}& 80.9\% & 92.9\% \\
\hline
\textbf{\textit{variation}}& 82.9\% & 87.5\% \\
\hline
\textbf{\textit{customText}}& 80.0\% & 91.2\% \\
\hline
\end{tabular}
\label{table3}
\end{center}
\end{table}

Beyond the fact that the results themselves can be considered as acceptable, the other behavior to be highlighted in Table~\ref{table3} is the greater classification capacity into \textit{standard clauses}, \textit{variations} and \textit{customized texts} of the NLP Engine in one document concerning another.

\subsection{Accuracy determination based on symbol comparison}
The second accuracy analysis involves establishing a comparison between the sub-articles populated in the output file concerning the sub-articles existing in the input file containing the real RAs. This comparison is performed at the symbol level, considering the order of appearance of each symbol. For that purpose, the text comparison tool Countwordsfree (\cite{countwordsfree}) is used by manually copying sub-article by sub-article. For each sub-article, the following values are determined:

\begin{enumerate}
\item Common percentage of symbols between compared sub-articles.
\item Difference percentage of symbols between compared sub-articles.
\item Common symbols between compared sub-articles.
\item Difference symbols between compared sub-articles.
\end{enumerate}

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.43\textwidth]{images/Proximus.png}}
\caption{Common percentage of symbols for Proximus Roaming Agreement.}
\label{fig3}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.43\textwidth]{images/Orange.png}}
\caption{NLP Engine overall architecture for Orange Roaming Agreement.}
\label{fig4}
\end{figure}

The illustrated radar charts allow us to analyze the results for each RA. The radar chart contains in the inner radius the common percentage of symbols between compared sub-articles and in the outer radius the identifier of the compared sub-article, i.e., the sub-articles contained in the output file. Fig.~\ref{fig3} shows a comparison in terms of the common percentage of symbols between compared sub-articles for the Proximus sample RA. The conclusion that can be reached by simple inspection is that only 11 of the 72, i.e.,  84.7\% of the analyzed sub-articles are affected, mostly by the introduction of undesired characters when the output file was populated.

Similarly, Fig.~\ref{fig4} shows a comparison in terms of the common percentage of symbols between the sub-articles compared for the Orange sample RA. In this case, only 7 of the 86, i.e., 91.9\% of the analyzed sub-articles are affected by the introduction of undesired characters when the output file is populated, thus improving the results obtained regarding the RA Proximus sample.

%\subsection{Analysis of the novelty of the contribution}

%\begin{table}[htbp]
%\caption{Analysis of the novelty of the contribution.}
%\begin{center}
%\begin{tabular}{|c|c|c|c|c|c|c|c|}
%\hline
%\multirow{2}{*}{\textbf{Field}} & %\multirow{2}{*}{\textbf{C}} & %\multicolumn{3}{c|}{\textbf{RA phase}} & %\multicolumn{3}{c|}{\textbf{NLP  Engine}} \\ %\cline{3-8} 
%&  & Dr & I & T & D & I & T \\ \hline
%\textbf{} & our contribution &  &  &  &  & &  \\ %\hline
%\multirow{3}{*}{\textbf{RA}} 
%& \cite{9369516} &  &  &  &  &  &  \\ \cline{2-8} 
%& \cite{9024541} &  &  &  &  &  &  \\ \cline{2-8} 
%& \cite{Huillet2020} &  &  &  &  &  &  \\ \hline
%\multirow{3}{*}{\textbf{NLP}} 
%& \cite{8487847} &  &  &  &  &  &  \\ \cline{2-8} 
%& \cite{9138070} &  &  &  &  &  &  \\ \cline{2-8} 
%& \cite{9104105} &  &  &  &  &  &  \\ \hline
%\end{tabular}
%\label{table3}
%\end{center}
%\end{table}